# global-syndemic

<!-- Quarto render -->

```{r}
#| label: setup
#| include: false

library(badger)
library(beepr)
library(groomr) # github.com/danielvartan/groomr
library(lubridate)
library(readr)
library(stringr)
```

<!-- badges: start -->
```{r}
#| echo: false
#| output: asis

library(badger)

cat(
  badger::badge_repostatus("wip"),
  badger::badge_custom(
    x = "CoMSES Network",
    y = "not published",
    color = "F5514D", # "1284C5"
    url = "https://www.comses.net/"
  ),
  badger::badge_custom(
    x = "OSF DOI",
    y = "10.17605/OSF.IO/852KG",
    color = "1284C5",
    url = "https://doi.org/10.17605/OSF.IO/852KG"
  ),
  # paste0(
  #   "[![",
  #   "FAIR checklist badge",
  #   "](",
  #   "https://img.shields.io/badge/",
  #   "fairsoftwarechecklist.net", "-",
  #   "-",
  #   "00a7d9",
  #   ")](",
  #   "https://fairsoftwarechecklist.net/v0.2?f=31&a=30112&i=32301&r=123",
  #   ")"
  # ),
  # paste0(
  #   "[![",
  #   "fair-software.eu",
  #   "](",
  #   "https://img.shields.io/badge/",
  #   "fair--software.eu", "-",
  #   "%E2%97%8F%20%20%E2%97%8F%20%20%E2%97%8F%20%20%E2%97%8F%20%20%E2%97%8F",
  #   "-",
  #   "green",
  #   ")](",
  #   "https://fair-software.eu",
  #   ")"
  # ),
  badger::badge_license(
    license = "GPLv3",
    color = "bd0000",
    url = "https://www.gnu.org/licenses/gpl-3.0"
  ),
  badger::badge_custom(
    x = "Contributor%20Covenant",
    y = "2.1",
    color = "4baaaa",
    url = "CODE_OF_CONDUCT.md"
  )
)
```
<!-- badges: end -->

## Overview

`Global Syndemic` is a [NetLogo](https://www.netlogo.org) model designed to simulate how climate change can affect the health and nutrition of children under 5 years old in Brazil. It is part of a larger project called: *Global syndemic: the impact of anthropogenic climate change on the health and nutrition of children under five years old attended by Brazil's public health system (SUS)*.

The model runs in paralel with the [`LogoClim`](https://github.com/sustentarea/logoclim) model, which provides climate data from [WorldClim 2.1](https://worldclim.org/), and the [`FoodClim`](https://github.com/sustentarea/foodclim) model, which simulates food yield responses to climate change.

> If you find this project useful, please consider giving it a star! &nbsp; [![GitHub repo stars](https://img.shields.io/github/stars/sustentarea/global-syndemic)](https://github.com/sustentarea/global-syndemic/)

## How to Use It

Refer to the [`LogoClim`](https://github.com/sustentarea/logoclim) and [`FoodClim`](https://github.com/sustentarea/foodclim) installation guides for detailed steps on installing the required dependencies.

Once both are installed, you can run the `Global Syndemic` model by specifying the path to your `LogoClim` and `FoodClim` installation in the `Global Syndemic` interface.

Refer to the `Info` tab in the model for additional details.

## How to Cite

If you use this model in your research, please cite it to acknowledge the effort invested in its development and maintenance. Your citation helps support the ongoing improvement of the model.

To cite `Global Syndemic` in publications please use the following format:

Vartanian, D., Garcia, L., & Carvalho, A. M. (2025). *Global syndemic: Simulating climate change effects on child nutrition in Brazil* [Computer software]. [https://doi.org/10.17605/OSF.IO/852KG](https://doi.org/10.17605/OSF.IO/852KG)

A BibTeX entry for LaTeX users is:

```latex
@Misc{vartanian2025,
  title = {Global syndemic: Simulating climate change effects on child nutrition in Brazil},
  author = {{Daniel Vartanian} and {Leandro Garcia} and {Aline Martins de Carvalho}},
  year = {2025},
  doi = {https://doi.org/10.17605/OSF.IO/852KG},
  note = {NetLogo model}
}
```

## How to Contribute

```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_custom( # Contributor Covenant
    "Contributor%20Covenant",
    "2.1",
    "4baaaa",
    "CODE_OF_CONDUCT.md"
  )
)
```

Contributions are welcome! Whether you want to report bugs, suggest features, or improve the code or documentation, your input is highly valued.

When contributing code, please follow the [tidy design principles](https://design.tidyverse.org/) and the [tidyverse style guide](https://style.tidyverse.org/) whenever possible.

## License

```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_license(
    license = "GPLv3",
    color = "bd0000",
    url = "https://www.gnu.org/licenses/gpl-3.0"
  )
)
```

```text
Copyright (C) 2025 Daniel Vartanian

Global Syndemic is free software: you can redistribute it and/or modify it
under the terms of the GNU General Public License as published by the Free
Software Foundation, either version 3 of the License, or (at your option) any
later version.

This program is distributed in the hope that it will be useful, but WITHOUT ANY
WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A
PARTICULAR PURPOSE. See the GNU General Public License for more details.

You should have received a copy of the GNU General Public License along with
this program. If not, see <https://www.gnu.org/licenses/>.
```

## Acknowledgments

We gratefully acknowledge [Stephen E. Fick](https://orcid.org/0000-0002-3548-6966), [Robert J. Hijmans](https://orcid.org/0000-0001-5872-2872), and the entire [WorldClim](https://worldclim.org/) team for their outstanding work in creating and maintaining the WorldClim datasets, which form the foundation of this project.

We thank the [Climatic Research Unit](https://www.uea.ac.uk/groups-and-centres/climatic-research-unit) at the [University of East Anglia](https://www.uea.ac.uk/) and the United Kingdom's [Met Office](https://www.metoffice.gov.uk/) for developing and providing access to the [CRU-TS-4.09](https://crudata.uea.ac.uk/cru/data/hrg/cru_ts_4.09/) dataset, a vital source of historical climate data.

We also acknowledge the World Climate Research Programme ([WCRP](https://www.wcrp-climate.org/)), its Working Group on Coupled Modelling, and the Coupled Model Intercomparison Project Phase 6 ([CMIP6](https://pcmdi.llnl.gov/CMIP6/)) for coordinating and advancing global climate model development.

We are grateful to the climate modeling groups for producing and sharing their model outputs, the Earth System Grid Federation ([ESGF](https://esgf.llnl.gov/)) for archiving and providing access to the data, and the many funding agencies that support CMIP6 and ESGF.

```{r}
#| echo: false
#| output: asis

htmltools::tags$table(
  htmltools::tags$tr(
    htmltools::tags$td(
      htmltools::tags$br(),
      htmltools::tags$br(),
      htmltools::tags$p(
        htmltools::tags$a(
          htmltools::tags$img(
            src = "images/sustentarea-logo.svg",
            width = 125
          ),
          href = "https://www.fsp.usp.br/sustentarea/"
        ),
        align = "center"
      ),
      htmltools::tags$br(),
      width = "30%"
    ),
    htmltools::tags$td(
      htmltools::tags$p(
        "This work was developed with support from the ",
        htmltools::tags$a(
          "Sustentarea",
          href = "https://www.fsp.usp.br/sustentarea/"
        ),
        " Research and Extension Center at the University of São Paulo (",
        htmltools::tags$a(
          "USP",
          href = "https://www5.usp.br/",
          .noWS = "outside"
        ),
        ")."
      ),
      width = "70%"
    )
  )
) |>
  knitr::raw_html()

htmltools::tags$table(
  htmltools::tags$tr(
    htmltools::tags$td(
      htmltools::tags$br(),
      htmltools::tags$p(
        htmltools::tags$a(
          htmltools::tags$img(
            src = "images/cnpq-logo.svg",
            width = 150
          ),
          href = "https://www.gov.br/cnpq/"
        ),
        align = "center"
      ),
      htmltools::tags$br(),
      width = "30%"
    ),
    htmltools::tags$td(
      htmltools::tags$p(
        "This work was supported by the Department of Science and ",
        "Technology of the Secretariat of Science, Technology, and Innovation ",
        "and of the Health Economic-Industrial Complex (",
        htmltools::tags$a(
          "SECTICS",
          href = "https://www.gov.br/saude/pt-br/composicao/sectics/",
          .noWS = "outside"
        ),
        ")  of the ",
        htmltools::tags$a(
          "Ministry of Health",
          href = "https://www.gov.br/saude/pt-br/composicao/sectics/",
          .noWS = "outside"
        ), " ",
        "of Brazil, and the National Council for Scientific and ",
        "Technological Development (",
        htmltools::tags$a(
          "CNPq",
          href = "https://www.gov.br/cnpq/",
          .noWS = "outside"
        ),
        ") (grant no. 444588/2023-0)."
      ),
      width = "70%"
    )
  )
) |>
  knitr::raw_html()
```
